How do families shape average neighbourhood incomes in Toronto?

### 2024-12-04 paper/paper.qmd
Added another section to the appendix to complete the paper and rendered the pdf

In addition to this, the potential for error that exists within the Neighbourhood Profiles data from the 2021 Canadian Census (as briefly mentioned in @sec-datameasure) can be another limitation that hints at a compelling future direction for this analysis. As the Neighbourhood Profiles data is the aggregation of individual family or household characteristics and experiences, it can provide a better understanding of the overall picture – that is, the general patterns and trends within the city of Toronto. However, it is limited in capturing unique or individual characteristics which can be lost during its process of aggregating data. Thus, with this overall understanding about the key drivers of average neighbourhood income, future research could be focused on individual neighbourhoods (the neighbourhoods with many 4-person families, for instance). These future studies could look at how key families interact with the local resources (e.g. grocery stores, health services, etc.) available to them to understand how they contribute to their neighbourhood’s average income. It may also be interesting to examine other characteristics (e.g. education attainment, ethnicity, etc.) that might explain why 2 and 4-person families are drivers of the average income for their neighbourhoods.

@@

>>>>

In addition to this, the potential for error that exists within the Neighbourhood Profiles data from the 2021 Canadian Census (as briefly mentioned in @sec-datameasure) can be another limitation that hints at a compelling future direction for this analysis. As the Neighbourhood Profiles data is the aggregation of individual family or household characteristics and experiences, it can provide a better understanding of the overall picture – that is, the general patterns and trends within the city of Toronto. However, it is limited in capturing unique or individual characteristics which can be lost during its process of aggregating data. Thus, with this overall understanding about the key drivers of average neighbourhood income, future research could be focused on individual neighbourhoods (the neighbourhoods with many 4-person families, for instance). These future studies could look at how key families interact with the local resources (e.g. grocery stores, health services, etc.) available to them to understand how they contribute to their neighbourhood’s average income. It may also be interesting to examine other characteristics (e.g. education attainment, ethnicity, etc.) that might explain why 2 and 4-person families are drivers of the average income for their neighbourhoods. A discussion of how this research might be conducted using observational data and surveys can be found in @sec-appendixobs.

@@

## Using observational data and surveys to better understand family characteristics at the neighbourhood level {#sec-appendixobs}

As discussed in @sec-discussion, a compelling future direction is to look at how key families interact with the local resources (e.g. grocery stores, health services, etc.) available to them to understand how they contribute to their neighbourhood’s average income. The following outlines a research plan for this direction.

To explore this new research direction, a participant observation and survey method can be used to recognize how families of different sizes interact with local resources and contribute to their neighbourhood’s average income. The participant observation will take place at a specific neighbourhood in Toronto and provide a clearer idea of how 2 or 4-person families spend their income (i.e. what neighbourhood resources they spend their income on). To compliment this, a survey will be distributed to all families in a particular neighbourhood to gain a better understanding of other characteristics such as education attainment or ethnicity families within that neighbourhood have. The survey can also highlight more about the reason behind family’s decisions and the behaviours observed.

Field notes that contain descriptions of family’s behaviour such as how often 4-person families visit a grocery store or a library will be taken and collected throughout this observation process. In terms of the survey, its structure and the kinds of questions it contains will depend on the neighbourhood. If the analysis is taking place at a neighbourhood with a higher number of 4-person families, then it is likely that there more families with children. So, questions could be tailored to their experiences such as, “To what extent are your expenses driven by your children?”. These kinds of open-ended questions will allow for a comparison with the observational data. If many families are coming to grocery stores with carts full of children’s clothes and snacks, and if many families are indicating that most of their expenses center around their children, this suggests that planning decisions that help support these families may also enhance a neighbourhood’s income position.

Thus, observational data and surveys can be helpful tools to find out more about individual family characteristics that may often be overlooked or lost when looking at aggregated census data or considering all Toronto neighbourhoods at once.
